Combinatorics of the geometry of Wilson loop diagrams

### 2019-02-18 General WLD combinatorics.tex
Inside/outside dfn version 2.2.2
made some changes to section 3.1... especially regards to notation

This in turn defines two sets of propagators, $\{\cP_{in}, \cP_{out}\}$, where (for $\bullet \in \{in, \; out\}$) \bas \cP_{\bullet} = \{ q \in (\cP \setminus p) | V(q) \cap V_{\bullet}(p)  \neq \emptyset \} \;.\eas The set $\cP_{in}$ (resp. $\cP_{out}$) is the set of propagators, not equal to $p$ \hlfix{whose support contains points in $V_{in}(p, <_*)$ (resp. $V_{in}(p, <_*)$), where $<_*$ refers to the appropriate ordering.}{Note, this means that paralell propagators are neither on the inside and outside of $p$. Is that okay?} If $W$ has no propagators crossing $p$ and no other propagator supported entirely on $V(p)$, then the set $\{\cP_{in}, p, \cP_{out}\}$ partitions $\cP$. Under these circumstances, the set $\cP_{in}$ defines the propagators with end points on one side of $p$, while $\cP_{out}$ defines the propagators with end points on the other side. In particular, for (weakly) admissible Wilson loop diagrams, the set $\{\cP_{in}, p, \cP_{out}\}$ is a partition for every propagators $p \in \cP$.

@@

\begin{dfn}\note{right now this doesn't depend on the order of $i$ and $j$, but that might matter later on?}
Let $p = (i,j) \in cP$ be a propagator in $W$.  Define the \textit{length} of $p$ to be 
\[\ell(p) = min\big\{|V_{in}(p)|,|V_{out}(p)|\big\} = min\big\{|[i+1,j]|,|[j+1,i]|\big\}.\]
\end{dfn}
In other words, $\ell(p)$ is the size of the smaller of the two sets $V_{in}(p)$ and $V_{out}(p)$ defined above.

\begin{rmk}\label{rem:props of length 2 and 3}
The following observations about propagators of short length in $W$ are easily verified:
\begin{enumerate}
\item If $p = (i,i+3)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.
\end{enumerate}
\end{rmk}

The following lemma establishes certain configurations of propagators that must exist in any admissible diagram with no non-supporting vertices (i.e. with $F(\emptyset) = \emptyset$). We make use of this result in several induction proofs below. Recall from Definition~\ref{admisdfn} that a {\em weakly admissible} diagram is one that satisfies the density and non-crossing conditions but does not require that $|V(\cP)| \geq |\cP| + 4$.

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible WLD with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 inside it and nothing else inside it.\label{item big and 2}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
\end{lem}

\begin{proof}
Suppose first that $W$ has a propagator of length $3$, say $p=(i, i+3)$.  By Remark~\ref{rem:props of length 2 and 3} and the fact that every vertex of $W$ supports at least one propagator, we have that $i+2$ supports exactly one propagator and this propagator must have length 2 by noncrossingness.  This gives us an instance of configuration~\ref{item big and 2} from the statement.

@@

Note that if $p$ is a propagator supported on $(i,j)$, then we can consider two obvious restrictions of $W$, namely $W_{[i,j+1]}$ and $W_{[j,i+1]}$.  Clearly all other propagators $q\neq p$ appear in exactly one of these subdiagrams, while $p$ appears in both.  In this case, we will say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$; the ordering of the start and end points allow us to identify one or the other region unambiguously.

@@

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $q_r$ bounds a region of $W$ which contains strictly fewer vertices than that bounded by $q_{r-1}$ and the restriction of $W$ to this region does not contain any of the propagators $p_1, \dots, p_r$.
\end{itemize}
Since $W$ contains finitely many vertices, this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+2)$ for its support.  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise. 
By assumption: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported.  Therefore there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

>>>>

This in turn defines two sets of propagators, $\{\cP_{in}, \cP_{out}\}$, where (for $\bullet \in \{in, \; out\}$) \bas \cP_{\bullet} = \{ q \in (\cP \setminus p) | V(q) \cap V_{\bullet}(p)  \neq \emptyset \} \;.\eas The set $\cP_{in}$ (resp. $\cP_{out}$) is the set of propagators, not equal to $p$ \hlfix{whose support contains points in $V_{in}(p, <_*)$ (resp. $V_{in}(p, <_*)$), where $<_*$ refers to the appropriate ordering.}{Note, this means that paralell propagators are neither on the inside and outside of $p$. Is that okay?} If $W$ has no propagators crossing $p$ and no other propagator supported entirely on $V(p)$, then the set $\{\cP_{in}, p, \cP_{out}\}$ partitions $\cP$. Under these circumstances, the set $\cP_{in}$ defines the propagators with end points on one side of $p$, while $\cP_{out}$ defines the propagators with end points on the other side. In particular, for (weakly) admissible Wilson loop diagrams, the set $\{\cP_{in}, p, \cP_{out}\}$ is a partition for every propagators $p \in \cP$. Recall from Definition~\ref{admisdfn} that a {\em weakly admissible} diagram is one that satisfies the density and non-crossing conditions but does not require that $|V(\cP)| \geq |\cP| + 4$.

@@

\begin{dfn}\note{right now this doesn't depend on the order of $i$ and $j$, but that might matter later on?}
Let $p = (i,j) \in cP$ be a propagator in $W$.  Define the \textit{length} of $p$ to be 
\[\ell(p) = min\big\{|V_{in}(p)|,|V_{out}(p)|\big\} +2 = min\big\{|[i+1,j]|,|[j+1,i]|\big\}.\]
\end{dfn}
In other words, $\ell(p)$ is the size of the smaller of the number of vertices that lie on either side of the progator $p$.

\begin{rmk}\label{rem:props of length 2 and 3} \sanote{change remark to "Configurations"?}
The following observations about propagators of short length in an (weakly) admissibly Wilson loop diagram, $W$, are easily verified:
\begin{enumerate}
\item If $p = (i,i+3)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.
\end{enumerate}
\end{rmk}

The following lemma establishes certain configurations of propagators that must exist in any admissible diagram with no non-supporting vertices (i.e. with $F(\emptyset) = \emptyset$). We make use of this result in several induction proofs below.

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible WLD with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 inside it and nothing else inside it.\label{item big and 2} \sanote{doesn't this depend on which side you call inside? Should we say ``on one side''?}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
\end{lem}

\begin{proof} \sanote{fixed misuse of the word support in lemma}
Suppose first that $W$ has a propagator of length $3$, say $p=(i, i+3)$.  By Remark~\ref{rem:props of length 2 and 3} and the fact that every vertex of $W$ supports at least one propagator, we have that $i+2$ supports exactly one propagator and this propagator must have length 2 by noncrossingness.  This gives us an instance of configuration~\ref{item big and 2} from the statement.

@@

Note that if $p = (i, j)$, then we can consider two obvious restrictions of $W$, namely $W_{[i,j+1]}$ and $W_{[j,i+1]}$.  Clearly all other propagators $q\neq p$ appear in exactly one of these subdiagrams, while $p$ appears in both.  In this case, we will say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$; the ordering of the start and end points allow us to identify one or the other region unambiguously. \sanote{see e-mail about contracted WLD}

@@

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $q_r$ bounds a region of $W$ which contains strictly fewer vertices than that bounded by $q_{r-1}$ and the restriction of $W$ to this region does not contain any of the propagators $p_1, \dots, p_r$. \sanote{so I understand, $q_r$ on the other side of $q_{r-1}$ than the $\{p_i\}$?} 
\end{itemize}
Since $W$ contains finitely many vertices, this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1 = (j_1,j_1+2)$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}).  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise. 
By assumption: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported.  Therefore there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

INSERT PICTURE HERE?
